---
title: "A New Robust Multivariate EWMA Dispersion Control Chart for Individual Observations"
author: "Jimoh O. Ajadi"
date: "15 April 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r install-packages, include=FALSE, eval=FALSE}
 install.packages("mvtnorm")
 install.packages("expm") 
 install.packages("ICSNP")
 install.packages("caret")
 install.packages("MVN")
```

```{r load-packages, include=FALSE, eval=TRUE, message=FALSE,warning=TRUE}
require(mvtnorm)
require(expm)
require(ICSNP)
require(caret)
require(MVN)
```

## Call for multivariate gamma function
multivarate gamme functions can be found in the file multvariate_gamma_function.R" 
```{r echo=TRUE}
source("multivarate_gamma_function.R")
```

# Comparison of two vectors
The functions returns the element-wise maximum or minimum values between two
vectors. The inputs for the functions are

* Sided: upper-sided or lower-sided monitoring
* V1: vector 1
* V2: vector 2

```{r load-input-data}
comparevec=function(sided,S1,S2){
  n=length(S1)
  P=rep(0,n)
  for(i in 1:n){
    if(sided=="upper"){
      P[i]=max(S1[i],S2[i])
    }
    else if(sided=="lower"){
      P[i]=min(S1[i],S2[i])
    }
    
  }
  return(P)
}

```

# Out-of-control model
The functions returns the Phase II random vector based on Phase I estimates. The inputs for the functions are

* rho: correlation between the quality characteristic 
* p: number of correlated characteristic

```{r plot case data}
outofcontrolmodel2=function(rho,p){
  S=matrix(0,ncol = p,nrow = p)
  for(i in 1:p){
    for(j in 1:p){
      S[i,j]=rho^(abs(j-i))
    }
  }
  S1=S
  return(S1)
}
```

# generate Phase I data
The functions returns the Phase I estimated parameters. The inputs for the functions are

* m: Phase I sample size
* mu: In-control process mean vector
* sig0: Incontrol Covariance Matrix
* p: number of correlated characteristic
* df: degree of freedom 
* Distrib: normal, t, or gamma distributions
* ChartType: Name of the chart
* rate: rate of gamma distributed data
* shape: shape of gamma distributed data
```{r}
fPhaseI=function(m,mu,sig0,df,p,Distrib,ChartType,rate,shape){
  # generate data based on type of data model Distrib
  if(Distrib=="normal"){
      x=rmvnorm(m,mu,sig0)
    }else if(Distrib=="t"){
      x=mvtnorm:::rmvt(m,sigma=((df-2)/df)*sig0, df=df)
    }else if(Distrib=="gamma"){
      x=rmvgamma(m, shape=shape, rate=rate, corr=sig0)-rep(shape,p)
  }
  
  #estimate Phase I paramters according to equations (3), (4) and (9)
  if(ChartType=="MNSE"||ChartType=="NonParametric"){
      res <- HR.Mest(x,maxiter = 1000)
      m1=res$center
      SIG=res$scatter
      SIGinv=solve(sqrtm(SIG))
  }else{
      m1=colMeans(x)
      SIG=cov(x)
      SIGinv=solve(sqrtm(SIG))
  }
  
  return(list(c(m1),SIG,SIGinv))
}

```

# generate Phase II data
The functions returns the Phase II random vector based on Phase I estimates. The inputs for the functions are

* mu: In-control process mean vector
* MU: Phase I mean vector, 
* sig0: Incontrol Covariance Matrix
* SIGinv: Inverse of Covariance Matrix
* df: degree of freedom 
* p: number of correlated characteristic
* Distrib: normal, t, or gamma distributions
* ChartType: Name of the chart
* rate: rate of gamma distributed data
* shape: shape of gamma distributed data
```{r Phase II data}

fPhaseII=function(mu,MU,sig0,SIGinv,df,p,Distrib,ChartType,rate, shape){
 
  # generate Phase II data based on type of data model Distrib
  if(Distrib=="normal"){
      x=rmvnorm(1,mu,sig0)
    }else if(Distrib=="t"){
      x=mvtnorm:::rmvt(1,sigma=((df-2)/df)*sig0, df=df)
    }else if(Distrib=="gamma"){
      x=rmvgamma(1, shape=shape, rate=rate, corr=sig0)-rep(shape,p)
  }
  
  #transform Phase II paramters according to equation (10)
  if(ChartType=="MNSE"){
      y=spatialSign((x-MU)%*%SIGinv)
  }else if(Distrib=="t"){
      y=t(SIGinv%*%t(x-MU))
  }else{
      y=t(SIGinv%*%(x-MU))
  }
  return(y)
}
```

# function to obtain the conditional ARL of the MNSE chart
The functions returns the conditional ARL of the MNSE chart based on Phase I estimates. The inputs for the functions are

* mu: In-control process mean vector
* MU: Phase I mean vector, 
* sig0: Incontrol Covariance Matrix
* SIGinv: Inverse of the square root of Phase I Covariance Matrix
* w: smoothing parameters
* L: control constant
* df: degree of freedom 
* p: number of correlated characteristic
* Distrib: normal, t, or gamma distributions
* ChartType: Name of the chart
* rate: rate of gamma distributed data
* shape: shape of gamma distributed data
```{r}
fNParametric=function(mu,MU,sig0,SIGinv,w,UCL,Distrib,ChartType,df,rate,shape){
  y=fPhaseII(mu,MU,sig0,SIGinv,df,p,Distrib,ChartType,rate, shape)
  S=w*t(y)%*%(y)+(1-w)*diag(p)/p
  Q=sqrt(((2-w)/w)*sum(diag((p*S-diag(p))%*%(p*S-diag(p)))))
  rl=1
  while(Q<UCL){
    rl=rl+1
    y=fPhaseII(mu,MU,sig0,SIGinv,df,p,Distrib,ChartType,rate, shape)
    S=w*t(y)%*%(y)+(1-w)*S
    Q=sqrt(((2-w)/w)*sum(diag((p*S-diag(p))%*%(p*S-diag(p)))))
  }
  return(rl)
}

```

# function to obtain conditional ARL of the MEWMS chart
The functions returns the conditional ARL of the MEWMS chart based on Phase I estimates. The inputs for the functions are

* mu: In-control process mean vector
* MU: Phase I mean vector, 
* sig0: Incontrol Covariance Matrix
* SIGinv: Inverse of the square root of Phase I Covariance Matrix
* w: smoothing parameters
* L: control constant
* df: degree of freedom 
* p: number of correlated characteristic
* Distrib: normal, t, or gamma distributions
* ChartType: Name of the chart
* rate: rate of gamma distributed data
* shape: shape of gamma distributed data
```{r}
fUCL = function(w,p,rl,L){
  ucl = p+L*sqrt(2*p*((w/(2-w))+((2-2*w)/(2-w))*(1-w)^(2*(rl-1))))
  return(ucl)
}

fLCL = function(w,p,rl,L){
  lcl = p-L*sqrt(2*p*((w/(2-w))+((2-2*w)/(2-w))*(1-w)^(2*(rl-1))))
  return(lcl)
}

fMEWMS=function(p,mu,MU,sig0,SIGinv,w,L,Distrib,ChartType,df,rate,shape){
  rl=1
  y=fPhaseII(mu,MU,sig0,SIGinv,df,p,Distrib,ChartType,rate, shape)
  S=w*t(y)%*%(y)+(1-w)*t(y)%*%(y)
  SS=sum(diag(S))
  UCL= fUCL(w,p,rl,L)
  LCL= fLCL(w,p,rl,L)
  while(SS>LCL & SS<UCL){
    rl=rl+1
    y=fPhaseII(mu,MU,sig0,SIGinv,df,p,Distrib,ChartType,rate, shape)
    S=w*t(y)%*%(y)+(1-w)*S
    if(rl<10000){
      SS=sum(diag(S))
      UCL=fUCL(w,p,rl,L)
      LCL=fLCL(w,p,rl,L)
    }else{
      SS=sum(diag(S))
      UCL=SS-1
      LCL=SS+1
    }
  }
  return(rl)
}

```

# fMaxMEWMV: The next functions returns the the conditional ARL of MaxMEWMV Chart.
The functions returns the conditional ARL of the MaxMEWMV chart based on Phase I estimates. The inputs for the functions are

* mu: In-control process mean vector
* MU: Phase I mean vector, 
* sig0: Incontrol Covariance Matrix
* SIGinv: Inverse of the square root of Phase I Covariance Matrix
* w: smoothing parameters
* L: control constant
* df: degree of freedom 
* p: number of correlated characteristic
* Distrib: normal, t, or gamma distributions
* ChartType: Name of the chart
* rate: rate of gamma distributed data
* shape: shape of gamma distributed data
```{r}
fMaxMEWMV=function(p,mu,MU,sig0,SIGinv,w,UCL,Distrib,ChartType,df,rate,shape){
  MaxDt=UCL-1;
  rl=0
  while(MaxDt<UCL){
    rl=rl+1
    y=fPhaseII(mu,MU,sig0,SIGinv,df,p,Distrib,ChartType,rate, shape)
    if(rl==1){
      S=w*t(y)%*%(y)+(1-w)*t(y)%*%(y)
    }
    else{
      S=w*t(y)%*%(y)+(1-w)*S
    }
    Dt1=sum((diag(S)-1)^2)
    Dt2=sum((S[lower.tri(S)])^2)
    suma2=w/(2-w)+2*((1-w)^(2*rl-1))/(2-w)
    suma4=w^4/(1-(1-w)^4)+((1-w)^(4*(rl-1))*(1-(1-w)^4-w^4))/(1-(1-w)^4)
    ExpD1=2*p*suma2
    ExpD2=(p*(p-1)/2)*suma2
    VarD1=p*(48*suma4+8*suma2^2)
    VarD2=p*(p-1)*(2*p-1)*suma4+p*(p-1)*suma2^2
    if(rl<10000){
      MaxDt=max((Dt1-ExpD1)/sqrt(VarD1),(Dt2-ExpD2)/sqrt(VarD2))
    }else{
      MaxDt=UCL+1
    }
  }
  return(rl)
}

```

# fUREWMV: The next functions returns the conditional ARL of REWMV Chart for monitoring the upper side
The functions returns the conditional ARL of the proposed chart based on Phase I estimates. The inputs for the functions are

* mu: In-control process mean vector
* MU: Phase I mean vector, 
* sig0: Incontrol Covariance Matrix
* SIGinv: Inverse of the square root of Phase I Covariance Matrix
* w: smoothing parameters
* UCL: upper control limit
* df: degree of freedom 
* p: number of correlated characteristic
* Distrib: normal, t, or gamma distributions
* ChartType: Name of the chart
* rate: rate of gamma distributed data
* shape: shape of gamma distributed data
```{r}
fUREWMV=function(p,mu,MU,sig0,SIGinv,w,UCL,Distrib,ChartType,df,rate,shape){
  rl=1
  RB=rep(-1.27,p)
  y=fPhaseII(mu,MU,sig0,SIGinv,df,p,Distrib,ChartType,rate, shape)
  S=w*log(y*y)+(1-w)*RB
  S1=comparevec(sided="upper",S,RB)
  SS=sum(S1)
  while(SS< UCL){
    rl=rl+1
    y=fPhaseII(mu,MU,sig0,SIGinv,df,p,Distrib,ChartType,rate, shape)
    S=w*log(y*y)+(1-w)*S
    S1=comparevec(sided="upper",S,RB)
    SS=sum(S1)
  }
  return(rl)
}

```

# fLREWMV: The next functions returns the conditional ARL of REWMV Chart for monitoring the lower side
The functions returns the conditional ARL of the proposed chart based on Phase I estimates. The inputs for the functions are

* mu: In-control process mean vector
* MU: Phase I mean vector, 
* sig0: Incontrol Covariance Matrix
* SIGinv: Inverse of the square root of Phase I Covariance Matrix
* w: smoothing parameters
* LCL: lower control limit
* df: degree of freedom 
* p: number of correlated characteristic
* Distrib: normal, t, or gamma distributions
* ChartType: Name of the chart
* rate: rate of gamma distributed data
* shape: shape of gamma distributed data
```{r}
fLREWMV=function(p,mu,MU,sig0,SIGinv,w,LCL,Distrib,ChartType,df,rate,shape){
  rl=1
  RB=rep(-1.27,p)
  y=fPhaseII(mu,MU,sig0,SIGinv,df,p,Distrib,ChartType,rate, shape)
  S=w*log(y*y)+(1-w)*RB
  S1=comparevec(sided="lower",S,RB)
  SS=sum(S1)
  while(SS> LCL){
    rl=rl+1
    y=fPhaseII(mu,MU,sig0,SIGinv,df,p,Distrib,ChartType,rate, shape)
    S=w*log(y*y)+(1-w)*S
    S1=comparevec(sided="lower",S,RB)
    SS=sum(S1)
  }
  return(rl)
}
```

# Function to compute the conditional ARL for each of the compared charts

```{r}
fARLComparedCharts=function(p,m,mu,sig.ic,sig.oc,simno,ChartType,Distrib,df,w,rate,shape,UCL){
  RL=rep(0,simno)
  estimate=fPhaseI(m,mu,sig.ic,df,p,Distrib,ChartType,rate,shape)
  MU=estimate[[1]] 
  SIGinv=estimate[[3]]
  
  for(i in 1:simno){
    if(ChartType=="MEWMS"){
      RL[i]=fMEWMS(p,mu,MU,sig.oc,SIGinv,w,UCL,Distrib,ChartType,df,rate,shape)
    }else if(ChartType=="MaxMEWMV"){
      RL[i]=fMaxMEWMV(p,mu,MU,sig.oc,SIGinv,w,UCL,Distrib,ChartType,df,rate,shape)
    }else if(ChartType=="UREWMV"){
      RL[i]=fUREWMV(p,mu,MU,sig.oc,SIGinv,w,UCL,Distrib,ChartType,df,rate,shape)
    }else if(ChartType=="LREWMV"){
      RL[i]=fLREWMV(p,mu,MU,sig.oc,SIGinv,w,UCL,Distrib,ChartType,df,rate,shape)
    }else if(ChartType=="MNSE"){
      RL[i]=fNParametric(mu,MU,sig.oc,SIGinv,w,UCL,Distrib,ChartType,df,rate,shape)
    }
  }
  ARL=mean(RL)
  if(any(RL==10000)){COUNT=1}else{COUNT=0}
  return(data.frame(ARL,COUNT))
}
```

# fPern is the function that return the percentage data that pass the Henze and Zirkler multivariate normality test at a significance level of 0.05

The inputs for the functions are

* mu: In-control process mean vector
* SIG: Incontrol Covariance Matrix
* df: degree of freedom 
* m: number of data
* Type: normal, t, or gamma distributions
* rate: rate of gamma distributed data
* shape: shape of gamma distributed data
* simno: number of simulation
```{r}
fPern<-function(mu,SIG,Type,shape,rate,df,m,simno,...){
  Normality=rep(0,simno)
  for(i in 1:simno){
    if(Type=='Normal'){ 
      x=rmvnorm(m,mu,SIG)
      result <- mvn(data = x, mvnTest = "hz")
    }
    else if(Type=='Gamma'){
      x=rmvgamma(m, shape=shape, rate=rate, corr=SIG)-rep(shape,p)
      result <- mvn(data = x, mvnTest = "hz")
    }
    else if(Type=='t'){
      x=mvtnorm:::rmvt(m,sigma=((df-2)/df)*SIG, df=df)
      result <- mvn(data = x, mvnTest = "hz")
    } 
    if(result$multivariateNormality$`p value`>0.05){
      Normality[i]=1
    }else{
      Normality[i]=0
    }
  }
  PerNormality=sum(Normality)*100/simno
  return(PerNormality)
}

```

#Table 2: Simulation for computing the percentage of Henze and Zirkler tests for multivariate normality test that do not reject at a significance level of 0.05.

```{r eval=FALSE, include=FALSE}
p=5;SIG=diag(p);mu=rep(0,p)
m1=c(30,50,70,100,150,200,500,1000,2000,5000)
OUTN=OUTT=OUTG=rep(0,length(m1))
for(i in 1:length(m1)){
  m=m1[i]
  OUTN[i]=fPern(mu,SIG,Type='Normal',m=m,simno=10000)
  OUTT[i]=fPern(mu,SIG,Type='t',df=15,m=m,simno=10000)
  OUTG[i]=fPern(mu,SIG,Type='Gamma',shape=1,rate=1,m=m,simno=10000)
}
output=data.frame(m1,OUTN,OUTT,OUTG)
output

```

# Table 3: Analysis of Phase I estimators on different distributions

```{r eval=FALSE, include=FALSE}
p=5
simno=10000
vM=10#c(10,25,50,75,100,125,150,200)
Nm = length(vM)
rate <- 1
shape<- 16
df=30
Distrib=c("normal","t","gamma")
mu=rep(0,p)
sig.ic=diag(p)
Estimator=c("NonParametric","Parametric")
EstimatorType=Estimator[1]
#Input parameter to change 
selectedDist = Distrib[1]
## SIMULATION
ARL=rep(0,simno)
VARARL=rep(0,Nm)
for(j in c(1:Nm)){
  m=vM[j]
  for(i in c(1:simno)){
    estimate=fPhaseI(m,mu,sig.ic,df,p,selectedDist,EstimatorType,rate,shape)
    ARL[i] <-sum(diag(estimate[[2]]))
  }
  VARARL[j]=var(ARL[1:i])
}  
output=data.frame(vM,VARARL)
output

```

```{r}
fucl=function(p,w,m,ChartType){
  if(p==2){
    if(w==0.1){
      if(m==50){
        UCLMaxMEWMV=5.2;UCLREWMV=-1.22;LCLREWMV=-4.108;UCLMEWMS=2.55;UCLNP=3.034
      }else if(m==200){
        UCLMaxMEWMV=4.83;UCLREWMV=-1.175;LCLREWMV=-4.23;UCLMEWMS=2.475;UCLNP=2.9
      }else if(m==2000){
        UCLMaxMEWMV=4.6;UCLREWMV=-1.166;LCLREWMV=-4.255;UCLMEWMS=2.475;UCLNP=2.83
      }  
    }else if(w==0.3){
      if(m==50){
        UCLMaxMEWMV=7.17;UCLREWMV=0.06;LCLREWMV=-6.45;UCLMEWMS=3.338;UCLNP=2.7
      }else if(m==200){
        UCLMaxMEWMV=6.907;UCLREWMV=0.076;LCLREWMV=-6.54;UCLMEWMS=3.38;UCLNP=2.66
      }else if(m==2000){
        UCLMaxMEWMV=6.74;UCLREWMV=0.08;LCLREWMV=-6.57;UCLMEWMS=3.38;UCLNP=2.637
      }   
    }
  }else if(p==5){
    if(w==0.1){
      if(m==50){
        UCLMaxMEWMV=7.9;UCLREWMV=-3.542;LCLREWMV=-8.84;UCLMEWMS=3.17;UCLNP=7.21
      }else if(m==200){
        UCLMaxMEWMV=5.055;UCLREWMV=-3.787;LCLREWMV=-9.173;UCLMEWMS=2.638;UCLNP=6.594
      }else if(m==2000){
        UCLMaxMEWMV=4.21;UCLREWMV=-3.874;LCLREWMV=-9.26;UCLMEWMS=2.47;UCLNP=6.332
      }   
    }else if(w==0.3){
      if(m==50){
        UCLMaxMEWMV=9.24;UCLREWMV=-1.235;LCLREWMV=-12.38;UCLMEWMS=3.908;UCLNP=6.675
      }else if(m==200){
        UCLMaxMEWMV=6.62;UCLREWMV=-1.57;LCLREWMV=-12.7;0.000752;UCLMEWMS=3.26;UCLNP=6.427
      }else if(m==2000){
        UCLMaxMEWMV=5.9;UCLREWMV=-1.68;LCLREWMV=-12.8;UCLMEWMS=3.076;UCLNP=6.36
      }   
    }
  }else if(p==25){
    if(w==0.1){
      if(m==50){
        UCLMaxMEWMV=84.27;UCLREWMV=-8.82;LCLREWMV=-34.72;UCLMEWMS=20.2;UCLNP=35.5
      }else if(m==200){
        UCLMaxMEWMV=8;UCLREWMV=-21.1;LCLREWMV=-38.97;UCLMEWMS=4.93;UCLNP=28.34
      }else if(m==2000){
        UCLMaxMEWMV=3.7;UCLREWMV=-23.4;LCLREWMV=-40.5;UCLMEWMS=2.55;UCLNP=26.92
      }   
    }else if(w==0.3){
      if(m==50){
        UCLMaxMEWMV=59;UCLREWMV=-0.7;LCLREWMV=-42.2;UCLMEWMS=17.12;UCLNP=31.12
      }else if(m==200){
        UCLMaxMEWMV=8;UCLREWMV=-13.89;LCLREWMV=-48;UCLMEWMS=4.67;UCLNP=27.75
      }else if(m==2000){
        UCLMaxMEWMV=4.7;UCLREWMV=-16.4;LCLREWMV=-49.68;UCLMEWMS=2.91;UCLNP=27.16
      }   
    }
  }
  if(ChartType=="MaxMEWMV"){UCL=UCLMaxMEWMV}
  else if(ChartType=="UREWMV"){UCL=UCLREWMV}
  else if(ChartType=="LREWMV"){UCL=LCLREWMV}
  else if(ChartType=="MEWMS"){UCL=UCLMEWMS}
  else if(ChartType=="MNSE"){UCL=UCLNP}
  return(UCL)
}
```

## Initialization

```{r echo=TRUE}
Distrib=c("normal","t","gamma")
## Initialization
p=2
l=0.3
simnoI=10
simnoII=10000
sig.ic=diag(p)
rho=0.3
# In-control model
sig.ic=diag(p)

# Out-of-control models partial shift
sigoc1=outofcontrolmodel2(rho=0,p)
sigoc1[1,1]=1.2

# Out-of-control models overall shift
sigoc2=0.2*outofcontrolmodel2(rho=0.3,p)

l.sig.oc=list(sig.ic,
              sigoc1,  #This can be replace with any partial shifts in Tables 5 and 7
              sigoc2)  #This can be replace with any overall shifts in Tables 4 and 6

Nshift=length(l.sig.oc)
vM=200#c(50,200,2000)
Nm = length(vM)
rate <- 1
shape<- 16
df=30
Charts=c("MEWMS","MaxMEWMV","MNSE","LREWMV","UREWMV")
mu=rep(0,p)

#Input parameter to change 
ChartType=Charts[1]
selectedDist = Distrib[3]

```

# Table 4-7: SIMULATIONS



```{r echo=TRUE}
for(j in c(1:Nm)){
  m=vM[j]
  AARL=rep(0,Nshift)
  COUNT=SDARL=AARL
  UCL=fucl(p,l,m,ChartType)

  for( s in c(1:Nshift)){
    count=ARL=rep(0,simnoII)  
    sig.oc = l.sig.oc[[s]]
    
    for(i in c(1:simnoII)){
      set.seed(i)
      arl <- fARLComparedCharts(p,m,mu,sig.ic,sig.oc,simnoI,ChartType,selectedDist,df,l,rate,shape,UCL)
      ARL[i]=arl$ARL;count[i]=arl$COUNT
    }
    AARL[s]=mean(ARL[1:i])
    SDARL[s]=sd(ARL[1:i])
    COUNT[s]=sum(count)
  }  
  output=data.frame(AARL,SDARL,COUNT)
  output
}
```

#Output
```{r echo=TRUE}
output
```

